### 2020-10-20 plan/tables/serenelli2017_seismo_table.tex

\begin{deluxetable}{rrrrr}
\tablecaption{An extract from Table 3 of \citetalias{serenelliFirstAPOKASCCatalog-2017} comprising measurements of the global asteroseismic parameters from the Sydney pipeline \citep{huberAutomatedExtractionOscillation-2009} with uncertainties from the spread of results from several other pipelines.\label{tab:seismo}}
\tablehead{\colhead{\textit{Kepler} ID} & \colhead{$\nu_{\max}$} & \colhead{$\sigma_{\nu_{\max}}$} & \colhead{$\Delta\nu$} & \colhead{$\sigma_{\Delta\nu}$}}
\startdata
2450729 & 1053.105 & 114.904 & 61.910 & 2.539 \\
2991448 & 1111.248 & 18.148 & 61.732 & 0.899 \\
3223000 & 2573.222 & 563.234 & 110.919 & 1.662 \\
3241581 & 2807.592 & 395.565 & 123.412 & 2.821 \\
3427720 & 2726.381 & 56.767 & 120.045 & 0.120
\enddata
\end{deluxetable}